\documentclass{article}

\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{extramarks}
\usepackage[usenames,dvipsnames]{color}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}

\usepackage{tikz}

\usetikzlibrary{automata,positioning,calc}

\topmargin=-0.45in
\evensidemargin=0in
\oddsidemargin=0in
\textwidth=6.5in
\textheight=9.0in
\headsep=0.25in

\linespread{1.1}

\pagestyle{fancy}
\lhead{\hmwkAuthorName}
\chead{\hmwkClass\ (\hmwkClassInstructor\ \hmwkClassTime): \hmwkTitle}
\rhead{\firstxmark}
\lfoot{\lastxmark}
\cfoot{}
\renewcommand\headrulewidth{0.4pt}
\renewcommand\footrulewidth{0.4pt}

\setlength\parindent{0pt}

\newcommand{\enterProblemHeader}[1]{
    \nobreak\extramarks{#1}{#1 continued on next page\ldots}\nobreak
    \nobreak\extramarks{#1 (continued)}{#1 continued on next page\ldots}\nobreak
}

\newcommand{\exitProblemHeader}[1]{
    \nobreak\extramarks{#1 (continued)}{#1 continued on next page\ldots}\nobreak
    \nobreak\extramarks{#1}{}\nobreak
}

\setcounter{secnumdepth}{0}
\newcounter{homeworkProblemCounter}

\newcommand{\homeworkProblemName}{}
\newenvironment{homeworkProblem}[1][Problem \arabic{homeworkProblemCounter}]{
    \stepcounter{homeworkProblemCounter}
    \renewcommand{\homeworkProblemName}{#1}
    \section{\homeworkProblemName}
    \enterProblemHeader{\homeworkProblemName}
}{
    \exitProblemHeader{\homeworkProblemName}
}

\newcommand{\problemAnswer}[1]{
\noindent\framebox[\columnwidth][c]{\begin{minipage}{0.98\columnwidth}#1\end{minipage}}
}

\newcommand{\homeworkSectionName}{}
\newenvironment{homeworkSection}[1]{
    \renewcommand{\homeworkSectionName}{#1}
    \subsection{\homeworkSectionName}
    \enterProblemHeader{\homeworkProblemName\ [\homeworkSectionName]}
}{
    \enterProblemHeader{\homeworkProblemName}
}

\newcommand{\hmwkTitle}{Homework\ \#7}
\newcommand{\hmwkDueDate}{April 1, 2013 at 11:59pm}
\newcommand{\hmwkClass}{CS331}
\newcommand{\hmwkClassTime}{9:00am}
\newcommand{\hmwkClassInstructor}{Professor Zhang}
\newcommand{\hmwkAuthorName}{Josh Davis}

\title{
    \vspace{2in}
    \textmd{\textbf{\hmwkClass:\ \hmwkTitle}}\\
    \normalsize\vspace{0.1in}\small{Due\ on\ \hmwkDueDate}\\
    \vspace{0.1in}\large{\textit{\hmwkClassInstructor\ \hmwkClassTime}}
    \vspace{3in}
}

\author{\textbf{\hmwkAuthorName}}
\date{}

\begin{document}

\maketitle

\pagebreak

\begin{homeworkProblem}
    Let \(K = \{\left<M\right> : \left<M\right> \notin L(M)\}\). Prove that \(K\) is not Turing
    recognizable.

    \begin{proof}
        To prove that \(K\) is not Turing recognizable, we will prove this by coming up with a contradiction.
        \\

        Assume that \(K\) is Turing recognizable and \(N\) is a Turing machine
        that recognizes it. \(K\) is the set of all encoded Turing machines
        that recognize the language that doesn't include their own encoding.
        \\

        Using this assumption, we can see that if we take the TM that
        recognizes \(K\), \(N\), we can see there are two cases in which this
        property of \(K\) still holds.
        \\

        \textbf{Case One}
        \\

        If \(N \in K\) then \(N \notin K\). This is because if it contains
        itself, then it must be removed because \(K\) cannot contain the
        encoding of itself.
        \\

        \textbf{Case Two}
        \\

        If \(N \notin K\) then \(N \in K\). This is because the encoding of
        \(N\) must be represented in \(K\) because it is the set of all TM that
        don't recognize themselves.
        \\

        Thus \(N \in K \iff N \notin K\).  This is clearly a contradiction and
        thus we have shown that \(K\) is not Turing recognizable.
    \end{proof}

\end{homeworkProblem}

\pagebreak

\begin{homeworkProblem}
    Prove the following statements:
    \begin{enumerate}
        \item \(L_1 \leq_{m} L_2\) and \(L_2 \leq_{m} L_3\) imply \(L_1 \leq_{m} L_3\).
        \item \(L_1 \leq_{T} L_2\) implies that \(\overline{L_1} \leq_{m} \overline{L_2}\).
    \end{enumerate}

    \textbf{Part One}

    \begin{proof}
        Let there be three TMs that decide the three languages. \(M_1\)
        recognizes \(L_1\), \(M_2\) recognizes \(L_2\), and \(M_3\) recognizes
        \(L_3\).
        \\

        According to many-one reduction, if one language reduces to another
        language, that means there is a computable function such that the
        function reduces the first language to the second.
        \\

        Therefore with \(L_1 \leq_{m} L_2\), there is a computable function,
        \(f_{12}\) where for every \(w\), \(w \in L_1 \iff f_{12}(w) \in L_2\).
        \\

        The same can be said for \(L_2 \leq_{m} L_3\), there is a computable
        function, \(f_{23}\) where for every \(w\), \(w \in L_2 \iff f_{23}(w) \in
        L_3\).
        \\

        Now to prove that \(L_1 \leq_{m} L_2\) and \(L_2 \leq_{m} L_3\) imply
        \(L_1 \leq_{m} L_3\), we will construct a new TM, \(N\), that
        recognizes \(L_1 \leq_{m} L_3\). We will construct \(N\) using the following
        computable function, \(f(w) = f_{23}(f_{12}(w))\):
        \\

        \(N = \)`` On input string \(w\):
        \begin{enumerate}
            \item Run \(M_1\) on \(w\), if it accepts, move on, else reject
            \item Compute \(f(w)\) using our new computable function
            \item Run \(M_3\) on the previously computed value, output whatever \(M_3\) outputs''
        \end{enumerate}

        Thus we have shown that when \(L_1\) reduces to \(L_2\) and \(L_2\) reduces to \(L_3\),
        we can construct a new many-reduction that reduces \(L_1\) to \(L_3\). Thus we have proved
        what we sought to prove and our proof is complete.
    \end{proof}

    \textbf{Part Two}

    \begin{proof}
        Since \(L_1\) is Turing reducible to \(L_2\), that means \(L_1\) is
        deciable relative to \(L_2\) and there is an oracle TM, \(M^1\), that can report
        whether or not any string \(w\) is a member of \(L_1\). Likewise there
        is another oracle TM, \(M^2\), that can report whether or not any
        string \(w\) is a member of \(L_2\).
        \\

        To show that this then implies that \(\overline{L_1} \leq_{m} \overline{L_2}\),
        we can construct a new oracle TM using these existing oracle TMs. Let
        our new oracle TM be \(N\) and defined as follows:
        \\

        \(N = \)`` On input string \(w\):
        \begin{enumerate}
            \item Query \(M^1\) with \(w\), if it answers \textbf{no}, continue, else reject
            \item Next query \(M^2\) with \(w\), if it answers \textbf{no} accept, else reject''
        \end{enumerate}

        This is valid because when we query the oracles, since we want the
        complement of the languages, we just answer the opposite of what the
        oracle tells us. Thus we can create a new oracle from the oracles for
        \(L_1\) and \(L_2\) that can show that \(L_1 \leq_{T} L_2\) implies
        that \(\overline{L_1} \leq_{m} \overline{L_2}\).
        \\

        Thus we have proven what we wanted to and our proof is complete.

    \end{proof}
\end{homeworkProblem}

\pagebreak

\begin{homeworkProblem}
    Prove that \(A_{TM} \not\leq_{m} \overline{A_{TM}}\), where \(A_{TM} =
    \{\left<M, w\right> : M \mbox{ is a TM and } M \mbox{ accepts } w \}\)

    \begin{proof}
        To prove this, we will do a proof by contradiction.
        \\

        Suppose that \(A_{TM} \leq_{m} \overline{A_{TM}}\), or \(A_{TM}\)
        reduces to \(\overline{A_{TM}}\). This means that there is a computable
        function, \(f\) that on every input \(w\), \(w
        \in A_{TM} \iff f(w) \in \overline{A_{TM}}\). Let \(M\) be the TM that
        recognizes \(A_{TM}\).
        \\

        According to the above, we can construct a new TM, \(N\) that
        recognizes \(\overline{A_{TM}}\). Let \(N\) be constructed as follows:

        \(N = \)`` On input string \(w\):
        \begin{enumerate}
            \item Compute \(f(w)\)
            \item Run \(M\) with \(w\), if \(M\) rejects, then accepts and if \(M\) accepts, then reject''
        \end{enumerate}

        As we can see, the only way to that \(A_{TM}\) reduces to
        \(\overline{A_{TM}}\) is if \(A_{TM}\) is decidable. Since we know that
        \(A_{TM}\) is not decidable, we have a contradiction.
        \\

        Since we have a contradiction \(A_{TM} \leq_{m}
        \overline{A_{TM}}\) cannot be true and we have concluded our proof.
    \end{proof}
\end{homeworkProblem}

\begin{homeworkProblem}
    Which of of the following PCP problems has a solution? Justify.

    \Large
    \begin{enumerate}
        \item \(\{\left[\frac{ab}{a}\right], \left[\frac{bb}{ab}\right],
                \left[\frac{aa}{ba}\right], \left[\frac{cc}{bc}\right],
                \left[\frac{aa}{ca}\right], \left[\frac{d}{cd}\right]\}\)
        \item \(\{\left[\frac{ab}{a}\right], \left[\frac{bb}{ab}\right],
                \left[\frac{aa}{ba}\right], \left[\frac{c}{bc}\right],
                \left[\frac{aa}{ca}\right], \left[\frac{d}{cd}\right]\}\)
    \end{enumerate}
    \normalsize

    \textbf{Part One}
    \\

    One possible solution is below:
    \[
        \left[\frac{ab}{a}\right] \left[\frac{cc}{bc}\right] \left[\frac{d}{cd}\right]
    \]

    This is a solution because if we read across the top, we get \(abccd\) and
    if we read across the bottom we get \(abccd\). Thus we have solved this
    given PCB problem.
    \\

    \textbf{Part Two}
    \\

    One possible solution is below:
    \[
        \left[\frac{ab}{a}\right] \left[\frac{aa}{ba}\right] \left[\frac{bb}{ab}\right] \left[\frac{c}{bc}\right]
    \]

    This is a solution because if we read across the top, we get \(abaabbc\) and
    if we read across the bottom we get \(abaabcc\). Thus we have solved this
    given PCB problem.
\end{homeworkProblem}

\pagebreak

\begin{homeworkProblem}
    Does the following PCP problem P have a solution?

    \[
        \{
            \left[\frac{a}{ab}\right],
            \left[\frac{b}{ccc}\right],
            \left[\frac{c}{b}\right],
            \left[\frac{c}{d}\right],
            \left[\frac{dddd}{}\right],
            \left[\frac{ddde}{e}\right],
        \}
    \]

    Yes, it has a solution. One possible solution is below:

    \[
        \left[\frac{a}{ab}\right]
        \left[\frac{b}{ccc}\right]
        \left[\frac{c}{d}\right]
        \left[\frac{c}{d}\right]
        \left[\frac{c}{d}\right]
        \left[\frac{ddde}{e}\right]
    \]

    This is a solution because if we read across the top, we get \(abcccddde\)
    and if we read across the bottom we get \(abcccddde\). Thus we have solved
    this given PCB problem.

\end{homeworkProblem}

\pagebreak

\begin{homeworkProblem}
    Show that \(BB(k)\) is not a computable function.

    \begin{proof}
        We will show that \(BB(k)\) is not a computable function by proof by
        contradiction.
        \\

        Assume that \(BB(k)\) is a computable function. Since \(BB(k)\) is a
        computable function, we can represent this function as a Turing
        machine. Let's name this TM, \(M_B\).
        \\

        \(M_B = \)`` On input \(k\) and \(w\), where \(k\) is a natural number and \(w\) is a TM encoding:
        \begin{enumerate}
            \item If \(w\) has k states, continue, else \(reject\).
            \item Run \(HALT_{TM}\) with input \(w\), if it accepts, continue, else reject.
        \end{enumerate}

        This works because we can reduce our problem of finding the TM that
        runs the longest to the halting problem.
        \\

        This is clearly a contradiction because we know that \(HALT_{TM}\) is
        undecidable thus by reduction, \(M_k\) is also undecideable
    \end{proof}
\end{homeworkProblem}

\end{document}
